Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-29 write_up/mixfishsim_ecolmod.tex
Response to JJP comments

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} \cdot \left(1 -
	\mathrm{e}^{-(F_{c,d} + M_{c,d})}\right) \cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch is the sum of
those across all fleets and vessels, $C_{c, d} =
\sum\limits_{fl=1}^{FL}\sum\limits_{v=1}^{V} E_{fl, v, c, d} \cdot Q_{fl} \cdot
D_{c, d}$ with $fl$ and $v$ the fleet and vessel respectively and
$E_{fl,v,c,d}$ and $Q_{fl}$ fishing effort and catchability of the gear, and
$D_{c,d}$ is the density of the population at the location fished. \\

@@

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c \in
\mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the joint
distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian with a
\textit{Matérn} covariance structure, where the correlation strength weakens
with distance. This enables us to model the spatial autocorrelation observed in
animal populations where density is more similar in nearby locations
\citep{Tobler1970, F.Dormann2007, Poos2007} and we change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field with an anisotropic pattern (to simulate a depth gradient) and
combine it with a temporally dynamic thermal tolerance field to imitate two key
drivers of population dynamics. Each population was initialised at a single
location, and subsequently moved across the entire space according to a
probabilistic distribution based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}^2$ is the squared index of
habitat suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see
below). \\

@@

\subsubsection{Decision about where to fish at the start of the trip}

Several studies \citep[for a review]{Girardin2016} have confirmed past activity
and past catch rates are strong predictors of fishing location choice. For this
reason, the fleet dynamics sub-model included a learning component, where a
vessel's initial fishing location in a trip was based on selecting from
previously successful fishing locations. This was achieved by calculating an
expected revenue based on the catches from locations fished in the preceding
trip as well as the same month periods in previous years and the travel costs
from the port to the fishing grounds. Then a vessel chooses randomly from the
top 70 \% of fishing events (defined as the `threshold') in terms of expected
profit within that season.

\subsubsection{Decision about where to fish within the trip}

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All population prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + Le \cdot \cos \left(\frac{\pi \cdot Br}{180}\right), \\
             & y1 + Le \cdot \sin \left(\frac{\pi \cdot Br}{180}\right) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $k$ the concentration parameter from the von
Mises distribution that we correlate with the revenue so that $k = (Rev + 1 /
RefRev) \cdot max_{k}$, where $max_{k}$ is the maximum concentration value,
$k$, and $RefRev$ is parametrised as for $\beta_{3}$ in the step length
function.

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/pop_dist}
	\caption{Spatial density (log abundance) for each of the four
		populations at four time steps. The darker the colour the
		greater the density of the population. Note that a diagonal
		anisotropic pattern representing a depth gradient can be
		clearly seen in populations 2 and 3. The concentrated spawning
		areas are also visible in the second row of the panels
		(t=18).}
	\label{fig:9}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Combined_Movement}
	\caption{(A) The fishing locations (points) and movements (lines) of a
		single vessel during a trip overlaid on the revenue of a
		fishing site (landings x price); (B) the fishing locations of
		the vessel over several trips (value field changes over the
		period so not shown). Note that movements are a mixture of
		correlated random walk and experience-based, and that the field
		is wrapped on a torus so that distances between opposite side
		of the spatial domain are considered next to each other. For
		these reasons there are moves larger than the step length in D;
		(C) the locations of multiple vessels from the same fleet
		overlaid on the value field, (D) the realised step distance and
		turning angles for a single vessel over the simulation.}
	\label{fig:mv}
\end{figure}

>>>>

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} \cdot \left(1 -
	\mathrm{e}^{-(F_{c,d} + M_{c,d})}\right) \cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch is the sum of
those across all fleets and vessels, $C_{c, d} =
\sum\limits_{fl=1}^{FL}\sum\limits_{v=1}^{V} E_{fl, v, c, d} \cdot Q_{fl} \cdot
D_{c, d}$ with $fl$ and $FL$ the fleet and total number of fleets, $v$ and $V$
the vessel and total number of vessels respectively and $E_{fl,v,c,d}$ and
$Q_{fl}$ fishing effort and catchability of the gear, and $D_{c,d}$ is the
density of the population at the location fished. \\

@@

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c
\in \mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the
joint distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian
with a \textit{Matérn} covariance structure, where the correlation strength
weakens with distance controlled by two parameters, with $\nu$ a scale
parameter in the units of distance and $\kappa$ a shape parameter which
determines the smoothness of the process. This enables us to model the spatial
autocorrelation observed in animal populations where density is more similar in
nearby locations \citep{Tobler1970, F.Dormann2007, Poos2007} and we change the
parameters to implement different spatial structures for the different
populations using the \textit{RandomFields} R package \citep{Schlater2015}. We
define a stationary habitat field with an anisotropic pattern (to simulate a
depth gradient) and combine it with a temporally dynamic thermal tolerance
field to imitate two key drivers of population dynamics without modelling the
processes explicitly. Each population was initialised at a single location, and
subsequently moved across the entire space according to a probabilistic
distribution based on habitat suitability (represented by the normalised values
from the GRFs), temperature and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}^2$ is the squared index of
habitat suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see
below). \\

@@

\subsubsection{Decision about where to fish at the start of a trip}

Several studies \citep[for a review see][]{Girardin2016} have confirmed past activity
and past catch rates are strong predictors of fishing location choice. For this
reason, the fleet dynamics sub-model included a learning component, where a
vessel's initial fishing location in a trip was based on selecting from
previously successful fishing locations. This was achieved by calculating an
expected revenue based on the catches from locations fished in the preceding
trip as well as the same month periods in previous years and the travel costs
from the port to the fishing grounds. Then a vessel chooses randomly from the
top 70 \% of fishing events (defined as the `threshold') in terms of expected
profit within that season.

\subsubsection{Decision about where to fish within a trip}

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All population prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
($x_{t}$,$y_{t}$) to ($x_{t+1}$, $y_{t+1}$) is defined by:
\begin{equation}
	\begin{split}
 (x_{t+1}, y_{t+1}) =  & x_{t} + Le \cdot \cos \left(\frac{\pi \cdot
		 Br_{t+1}}{180}\right), \\ & y_{t} + Le \cdot \sin
 \left(\frac{\pi \cdot Br_{t+1}}{180}\right) \\	
 when  \hspace{0.5cm}     & Br_{t} < 180, Br_{t+1} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t} > 180, Br_{t+1} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $Br_t$ is the bearing at time $t$, $k$ the
concentration parameter from the von Mises distribution that we correlate with
the revenue so that $k = (Rev + 1 / RefRev) \cdot max_{k}$, where $max_{k}$ is
the maximum concentration value, $k$, and $RefRev$ is parametrised as for
$\beta_{3}$ in the step length function.

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/pop_dist}
	\caption{Spatial density (log abundance) for each of the four
		populations at four time steps. The darker the colour the
		greater the density of the population. Note that a diagonal
		anisotropic pattern (mimicking a depth gradient) can be
		clearly seen in populations 2 and 3. The concentrated spawning
		areas are also visible in the second row of the panels
		(t=18).}
	\label{fig:9}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Combined_Movement}
	\caption{(A) The fishing locations (points) and movements (lines) of a
		single vessel during a trip overlaid on the revenue of a
		fishing site (landings x price); (B) the fishing locations of
		the vessel over several trips (value field changes over the
		period so not shown). Note that movements are a mixture of
		correlated random walk (solid lines) and experience-based
		(dashed lines), and that the field is wrapped on a torus so
		that opposite sides of the spatial domain are considered next
		door; (C) the locations of multiple vessels from the
		same fleet overlaid on the value field, (D) the realised step
		distance and turning angles for a single vessel over the
		simulation.}
	\label{fig:mv}
\end{figure}

### 2019-09-29 write_up/parameter_tables.tex
Response to JJP comments

\begin{table}[!ht]
	\centering
	\caption{Description of variables for population movement sub-module.}
	\begin{tabular}{p{1.5cm} p{8cm} p{1.5cm}}
		\toprule
		Variable & Meaning & Units \\
		\hline
		\multicolumn{3}{l}{\textit{Thermal tolerance}} \\
		\hline
		$T_{c,wk}$ & Temperature for cell in week & $^\circ$C \\
		$\mu_{p}$ & Mean of the thermal tolerance for population & $^\circ$C \\
		$\sigma^{2}_{p}$ & Standard deviation of thermal tolerance for
		the population & $^\circ$C \\
		\hline
		\multicolumn{3}{l}{\textit{Population movement model}} \\
		\hline
		$\lambda$ & Decay rate for population movement & - \\
		$Hab_{c,p}$ & Habitat suitability for cell $c$ and
		population $p$ & - \\
		$Tol_{c,wk,p}$ & Thermal tolerance for in cell $c$ at week $wk$
		for population $p$ & - \\
		$d_{I,J}$ & Euclidean distance between cell $I$ and cell $J$ & - \\
		\bottomrule
		\end{tabular}
	\end{table}

\begin{table}[!ht]
	\centering
	\caption{Description of variables for fleet dynamics sub-module.}
	\begin{tabular}{p{1.5cm} p{8cm} p{1.5cm}}
		\toprule
		Variable & Meaning & Units \\
		\hline
		$Rev$ & Revenue from fishing tow & \euro \\
		$L_{p}$ & Landings of population $p$ & kg \\
		$Pr_{p}$ & Average price of population $p$ & \euro
		\hspace{0.1cm} kg$^{-1}$ \\
		Le & Step length for vessel & - \\
		Br & Bearing & degrees \\
		$k$ & Concentration parameter for von mises distribution & - \\
		$\beta_{1}$ & shape parameter for step function & - \\
		$\beta_{2}$ & shape parameter for step function & - \\
		$\beta_{3}$ & shape parameter for step function & - \\
		\bottomrule
	\end{tabular}
\end{table}

@@

\begin{table}[!ht]
\caption{Fleet dynamics parameter setting.}
\begin{tabular}{ p{4cm } p{1.2cm} p{1.2cm} p{1.2cm} p{1.2cm} p{1.2cm}}
	\hline
	Parameter & Fleet 1 & Fleet 2 & Fleet 3 & Fleet 4 & Fleet 5 \\
	\hline
	Targeting preferences & pop 2/4 & pop 1/3 & - & pop 4 & pop
	2/3 \\
	\hline
	Price $Pr_{p}1$ = 100 & & & & & \\
	Price $Pr_{p}2$ = 200 & & & & & \\
	Price $Pr_{p}3$ = 350 & & & & & \\
	Price $Pr_{p}4$ = 600 & & & & & \\
	$Q_{p}$ & 0.01 & 0.02 & 0.02 & 0.01 & 0.01 \\
	$Q_{p}$ & 0.02 & 0.01 & 0.02 & 0.01 & 0.03\\
	$Q_{p}$ & 0.01 & 0.02 & 0.02 & 0.01 & 0.02 \\
	$Q_{p}$ & 0.02 & 0.01 & 0.02 & 0.05 & 0.01 \\
	\hline
	Exploitation dynamics &&&& \\
	\hline
	step function $\beta_1$ & 1 & 2 & 1 & 2 & 3 \\
	step function $\beta_2$ & 10 & 15 & 8 & 12 & 7 \\
	step function $\beta_3$ & Q90 & Q90 & Q85 & Q90 & Q80 \\
	step function $rate$ & 20 & 30 & 25 & 35 & 20 \\
	Past Knowledge = TRUE & & & & & \\
	Threshold & 0.7 & 0.7 & 0.7 & 0.7 & 0.7 \\
	Fuel Cost & 3 & 2 & 5 & 2 & 1 \\
	\hline
\end{tabular}
\label{tab:2}
\end{table}

>>>>

\begin{table}[!ht]
	\centering
	\caption{Description of variables for population movement sub-module.}
	\begin{tabular}{p{1.5cm} p{8cm} p{1.5cm}}
		\toprule
		Variable & Meaning & Units \\
		\hline
		\multicolumn{3}{l}{\textit{Thermal tolerance}} \\
		\hline
		$T_{c,wk}$ & Temperature for cell $c$ in week $wk$ & $^\circ$C \\
		$\mu_{p}$ & Mean of the thermal tolerance for population $p$ & $^\circ$C \\
		$\sigma^{2}_{p}$ & Standard deviation of thermal tolerance for
		population $p$ & $^\circ$C \\
		\hline
		\multicolumn{3}{l}{\textit{Population movement model}} \\
		\hline
		$\lambda$ & Decay rate for population movement & - \\
		$Hab_{c,p}$ & Habitat suitability for cell $c$ and
		population $p$ & - \\
		$Tol_{c,wk,p}$ & Thermal tolerance for in cell $c$ at week $wk$
		for population $p$ & - \\
		$d_{I,J}$ & Euclidean distance between cell $I$ and cell $J$ & - \\
		\bottomrule
		\end{tabular}
	\end{table}

\begin{table}[!ht]
	\centering
	\caption{Description of variables for fleet dynamics sub-module.}
	\begin{tabular}{p{1.5cm} p{8cm} p{1.5cm}}
		\toprule
		Variable & Meaning & Units \\
		\hline
		$Rev$ & Revenue from fishing tow & \euro \\
		$L_{p}$ & Landings of population $p$ & kg \\
		$Pr_{p}$ & Average price of population $p$ & \euro
		\hspace{0.1cm} kg$^{-1}$ \\
		Le & Step length for vessel & - \\
		$Br$ & Bearing & degrees \\
		$k$ & Concentration parameter for von mises distribution & - \\
		$\beta_{1}$ & shape parameter for step function & - \\
		$\beta_{2}$ & shape parameter for step function & - \\
		$\beta_{3}$ & shape parameter for step function & - \\
		\bottomrule
	\end{tabular}
\end{table}

@@

\begin{table}[!ht]
\caption{Fleet dynamics parameter setting.}
\begin{tabular}{ p{4cm } p{1.2cm} p{1.2cm} p{1.2cm} p{1.2cm} p{1.2cm}}
	\hline
	Parameter & Fleet 1 & Fleet 2 & Fleet 3 & Fleet 4 & Fleet 5 \\
	\hline
	Targeting preferences & pop 2/4 & pop 1/3 & - & pop 4 & pop
	2/3 \\
	\hline
	Price $Pr_{p}1$ = 100 & & & & & \\
	Price $Pr_{p}2$ = 200 & & & & & \\
	Price $Pr_{p}3$ = 350 & & & & & \\
	Price $Pr_{p}4$ = 600 & & & & & \\
	$Q_{p}$ & 0.01 & 0.02 & 0.02 & 0.01 & 0.01 \\
	$Q_{p}$ & 0.02 & 0.01 & 0.02 & 0.01 & 0.03\\
	$Q_{p}$ & 0.01 & 0.02 & 0.02 & 0.01 & 0.02 \\
	$Q_{p}$ & 0.02 & 0.01 & 0.02 & 0.05 & 0.01 \\
	\hline
	Exploitation dynamics &&&& \\
	\hline
	step function $\beta_1$ & 1 & 2 & 1 & 2 & 3 \\
	step function $\beta_2$ & 10 & 15 & 8 & 12 & 7 \\
	step function $\beta_3$, the landings value $nth$ quantile &
	90 & 90 & 85 & 90 & 80 \\
	step function $rate$ & 20 & 30 & 25 & 35 & 20 \\
	Past Knowledge = TRUE & & & & & \\
	Threshold & 0.7 & 0.7 & 0.7 & 0.7 & 0.7 \\
	Fuel Cost & 3 & 2 & 5 & 2 & 1 \\
	\hline
\end{tabular}
\label{tab:2}
\end{table}
